---
title: "Writing Reports with R Markdown"
author: "Felicia Bisnath"
date: "2022-10-21"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(knitr)
```
This report was prepared for the UN. It analyses the relationship between a country's GDP, life expectancy, and CO2 emissions. Our goal was to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesise that both life expectancy and CO2 emissions will increase with a country's GDP. 

``` {r data, include=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv")) #here figures out where the root is for the project

```

```{r gdp_lifeexp_1997, echo=FALSE}
ggplot(gapminder_1997, aes(x=gdpPercap, y=lifeExp, color=continent, size=pop/1000000)) +
  geom_point() +
  labs(x="GDP (per capita)",
       y="Life Expectancy",
       title="Do people in wealthier countries live longer?",
       size="Population (in millions)",
       color="Continent")
```

``` {r summary_data, include=FALSE}
n_countries <- gapminder_1997 %>%
  select(country) %>%
  n_distinct()

```

```{r minimum_gdp, include=FALSE}
min_gdp <- gapminder_1997 %>%
  summarise(min=round(min(gdpPercap)))

```

``` {r maximum_gdp, include=FALSE}
max_gdp <- gapminder_1997 %>%
  summarise(max=round(max(gdpPercap)))

```

The above plot shows the relationship between the GDP per capita and life expectancy for a total of `r n_countries` countries. Economic wealth ranged from a minimum of $ `r min_gdp` to a maximum of `r max_gdp` per capita.

|Summary of Data|
|-------|-------|
|Number of countries| `r n_countries`|
|Minimum GDP per cap | `r min_gdp`|
|Maximum GDP per cap | `r max_gdp`|

```{r tables, include=FALSE}
gapminder_1997 %>%
  filter(continent == "Oceania") %>%
  kable()

```

# Lists
## Unordered lists
* **R**
* _geom_
* **_statistics in R_**

## Ordered lists
1. **R**
2. _geom_
3. **_statistics in R_**
4. [Cheatsheets](www.rstudio.com/resources/cheatsheets)
5. [Markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)


1. list1
   - one
   - two
2. list2

```